# The Rectome Knowledgebase

- https://reactome.org
- opens source, open access, manually curated and peer-rewieved pathway database
- founded: 2003
- Lincoln Stein, Peter D’Eustachio, Henning Hermjakob, Guanming Wu
- human only
- contain: signal transduction, transport, DNA replication, protein synthesis, and intermediary metabolism
- single consistant data model
- can run in web browser form also!!
- https://reactome.org/about/statistics -statistics
- relased quarterly


## Reactom on Bioconductor
- bioconductor package: https://www.bioconductor.org/packages/release/bioc/html/ReactomePA.html
- publication: https://www.ncbi.nlm.nih.gov/pubmed/26661513

#### 1. Install package: 
```{r, include = F}
library(tidyverse)
library(readxl)
library(ReactomePA)
library(biomaRt)
```

```{r, eval = F}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ReactomePA")
```

#### 2. Read data:
```{r}
univ <- read_excel('./data/mRNA-seq_DE_table_two_sample_wt_vs_ko/all_genes.xlsx') %>% pull(ENSG)
targ <- read_excel('./data/mRNA-seq_DE_table_two_sample_wt_vs_ko/de_genes.xlsx') %>% pull(ENSG)
```

#### 3. Inspect data:
```{r}
length(univ)
univ[1: 6]
length(targ)
targ[1:6]
```

#### 4. Transltate Ensembl to Entrez by BioMart
```{r}
mart

getBM(
  attributes = c(
    
  ),
  
  mart = mart
    
  )
)
```


